---
title: "EDA on Vegas Strip dataset"
author: "Arun, Bronwyn, Manish"
date: "1/17/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```


```{r}

# loading required libraries
library(tidyverse)
library(caret)
library(readr)
library(ggplot2)
library(GGally)
library(repr)
library(gridExtra)



```

## Introduction and train/test split

In this document we perform exploratory data analysis on `Vegas Strip` dataaset to see if there is any relationship between different amenities which hotels provide and the score which guests give to those hotels on the strip. This exploratory analysis will then be extended to see if a predictive regression model can be developed to estimate user ratings based on hotel amenities.

Before doing any EDA splitting dataset into training and test set. 

```{r}
# reading dataset 
vegas_data <- read_csv2("../data/vegas_data.csv")

#splitting into training test data set  based on hotel names
set.seed(787)
train_index <- caret::createDataPartition(vegas_data$`Hotel name`,  p = 0.75, list = FALSE)
training_set <- vegas_data[train_index,]
test_set <- vegas_data[- train_index,]

dim(training_set)
dim(test_set)
```


### EDA 1:- Cleaning and wrangling data


Checking top rows and column names of the dataset. 


```{r}

head(training_set)

colnames(training_set)
```


The dataset has 20 columns, out of which 19 are features and `Score` is response variable for which relationship based on estimators need to be determined. 

The column/feature names are clearly not tidyverse compliant and have white spaces which are a bit difficult to deal with during data analysis. 

```{r}
#changing column names

colnames(training_set) <- c("user_country", "num_reviews", "num_hotel_reviews", "helpful_votes", "score", "stay_period", "traveller_type", "pool", "gym", "tennis_court", "spa", "casino", "free_internet", "hotel_name", "hotel_stars", "rooms", "user_continent", "member_years", "review_month", "review_weekday")

head(training_set)

```


Checking for any NAs in the dataset. 


```{r}

any(is.na(training_set))

```


This is good as we do not have any NAs in the dataset.


Checking datatype for individual columns in dataset.

```{r}

str(training_set)

# checking summary info for the training dataframe

summary(training_set)
```


Clearly based on summary statistics, the column `member_years` has negative values which seems incorrect as number of years of membership of a customer cannot be negative. 


```{r}
sort(training_set$member_years)[1:10]

```

Replacing the minimum value by 0 as it logically makes sense to do so.


```{r}
training_set[training_set$member_years== -1806,"member_years" ] = 0 

min(training_set$member_years)
```




### EDA 2 :- Exploring relationship b/w features and hotel score


Now we have clean data with no NA values, we can start inspecting individual features to see if we can identify any relationship between target score and the features.


```{r}
str(training_set)

```

Clearly the dataset has two kind of features.

- **User specific** : user_country, member_years, user_continent, traveller_type etc.
- **Hotel Specific** : rooms, free_internet, hotel_stars etc. 


And within each category we have both categorical and numerical features. 


We can first re-arrange columns for better interpretation and **remove hotel name** column as we do not want it to be one of the  features. 


```{r}
training_set <- training_set[c("user_country", "user_continent", "traveller_type",  "stay_period","review_month", "review_weekday", "member_years", "num_reviews", "helpful_votes", "pool", "gym", "tennis_court", "spa", "casino", "free_internet", "hotel_stars", "rooms", "num_hotel_reviews","score")]

head(training_set)

```


Creating vectors for different categories of features. 


```{r}

user_specific_features <- c("user_country", "user_continent", "traveller_type",  "stay_period","review_month", "review_weekday", "member_years", "num_reviews", "helpful_votes")

hotel_specific_features <- c("pool", "gym", "tennis_court", "spa", "casino", "free_internet", "hotel_stars", "rooms", "num_hotel_reviews")


categorical_features <- c("user_country", "user_continent", "traveller_type",  "stay_period","review_month", "review_weekday", "pool", "gym", "tennis_court", "spa", "casino", "free_internet", "hotel_stars")

numerical_features <- c( "member_years", "num_reviews", "helpful_votes",  "rooms", "num_hotel_reviews")

```

We already explored the numerical features and checked min/max values using `summary()` statistics earlier. Lets try exploring the `categorical features`.


```{r}
for (category in categorical_features){
  print(category)
  print(table(training_set[category]))
  cat("\n")
  cat("\n")
}


```

Interestingly, to our surprise there are  a few hotels on Vegas Strip which do not have casino. 

Also based on initial exploration of individual groups in each category; we are suspicious that users' `countries` should not have an impact on the hotel score. Also there were more visitors from some of the countries(USA, Canada, Australia) as compared to others(France etc.). 
Plus we may have a new user from different country so our model will more potentially fail to generalize and it might be a good idea to use `user_continents` instead. We will however do some more visualization analysis to confirm same. 


#### Inspecting Score's relationship with categorical features


Creating a function to plot categorical features against score

```{r}

# function to create plot for categorical features


plot_categorical <- function(df, col_name_x, col_name_y = "score"){
  
  p <-  ggplot(df, aes_string(x = col_name_x, y  = col_name_y))+
    geom_bar(stat = "summary", fun.y = "mean", fill = "royalblue3")+
    geom_jitter()+
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5),panel.background = element_rect(fill =   "lavender"))+
    ggtitle(paste("Average ", col_name_y," vs ", col_name_x))
  p
}

# list to hold plots

cat_plots <- vector('list', length(categorical_features))

cat_plots <- lapply(categorical_features, plot_categorical, df = training_set)

options(repr.plot.height = 10, repr.plot.width = 8)

grid.arrange(cat_plots[[1]], cat_plots[[2]], cat_plots[[3]],  layout_matrix = rbind(c(1,1),c(2,3)))
grid.arrange(cat_plots[[4]], cat_plots[[5]], cat_plots[[6]], cat_plots[[7]])
grid.arrange(cat_plots[[8]], cat_plots[[9]], cat_plots[[10]], cat_plots[[11]])
grid.arrange(cat_plots[[12]], cat_plots[[13]])
```


##### Observations

- Clearly based on the plots, some of the categorical features have signiciant impact on average user scores for hotels while others have minimal impact. 

- As analyzed earlier, based on plots of `score vs country` and `score vs continents`; it makes sense to use continents to study the relationship based on continents as compared to countries. This will help in better generalization of model.

- Most of the `Traveller types` seem to give similar averge hotel ratings with an exception of `Solo` travelers who rated hotels a bit lower. The travellers who visited in groups of friends rated the hotels higher as compared to other travel groups.


- Features `stay_period` and `review_weekday` seems to have same similar average scores for all stay periods and review days respectively. The average scores however sligtly seems to dip for mid-week reviews(Wednesday). We might need to see if these features can be removed from the model altogether to analyze/predict average scores.

- Amenities like `gym`,`tennis_court`, and `spa` seems to have slight positive impact on hotel scores. We however might need to check any collinearity within these features. Having a `pool` however improves hotel scores signiciantly. 

- For each `review_month`, there were small variations in score; it will however be a good idea to reduce the number of categories to four quarters rather than 12 months. 

- Providing visitors with `free internet` seems to improve hotel scores signicantly. 

- Although there is a small sample of hotels with no `casino` but those seems to have been rated better than the ones with casino. We will need to check if this is just due to class imbalance. 

- Different `Hotel stars` show variations in user scores. The trend needs to be investigated though.



Performing some wrangling based on the observations to further polish the dataset. 

```{r}
# removing user_countries column

training_set <- training_set[, -c(1)]

# Re-categorizing review months as quarters of a year rather than individual months. 
training_set$review_month <- factor(training_set$review_month)

levels(training_set$review_month) = list("Q1" = c("January", "February", "March"),
                                       "Q2" = c("April", "May", "June"),
                                       "Q3" = c("July", "August", "September"),
                                       "Q4" = c("October", "November", "December"))


# Re-categorizing review_weekday as weekends and weekdays rather than days of the week. 
# including Friday in weekend category

training_set$review_weekday <- factor(training_set$review_weekday)
levels(training_set$review_weekday) <- list("Weekday" = c("Monday", "Tuesday", "Wednesday", "Thursday"),
                                            "Weekend" = c("Friday","Saturday", "Sunday"))


# renaming these columns 

colnames(training_set)[4:5] <- c("review_quarter", "review_day")

# updating categorical feature vector 
categorical_features <- categorical_features[-1]
categorical_features[4:5] <- c("review_quarter", "review_day")

# updating user_specific feature vector
user_specific_features <- user_specific_features[-1]
user_specific_features[4:5] <- c("review_quarter", "review_day")




```


Visualizing updated relationship plots for these variables.



```{r}
# list to hold plots

cat_plots <- vector('list', length(categorical_features))

cat_plots <- lapply(categorical_features, plot_categorical, df = training_set)

grid.arrange(cat_plots[[4]], cat_plots[[5]])

```



#### Inspecting relationship between hotel scores and numeric features